---
title: "R Notebook"
output: html_notebook
---

Purpose: compare frequentist confidence intervals to Bayesian credible interval.

First generate a population to work with:

```{r}
n <- 1e5
r <- 0.7
pop <- data.frame(c(1:n), 0)
colnames(pop) <- c("id", "belief")
pop[,'belief'][sample(c(1:n), r*n)] <- 1
head(pop)
```

Now we take a sample and calculate a frequentist confidence interval. To do this we will use MLE. From a frequentist perspective, how does this work?

The frequentist believes that the parameter in question is fixed but unknown. This means we can't associate any probability distribution with it. From Bolstad:

>The only probability considered is the random sample of size $n$ given the parameter.

From this perspective, frequentist statistics is concerned with how statistics taken from a single random sample vary across all random samples. 

Given this random sample we want to make an estimate of the unknown parameter $\theta$. We can use the sample proportion, which turns out to be the MLE.

We take a sample of size $n$ from the population and calculate a statistic from that sample

```{r}
samp <- sample(pop[,"belief"], 100)
sampParams <- list(
  alpha = sum(samp),
  beta = length(samp) - sum(samp)
)

sampParams["est"] <- sampParams$alpha / (sampParams$alpha + sampParams$beta)
sampParams["interval"] <- qbeta(c(0.025, 0.975), sampParams$alpha, sampParams$beta)

with(sampParams, {
  curve(dbeta(x, alpha, beta), 0, 1)
  abline(v=est, col='blue')
  abline(v=r, col='red')
})
```

